\documentclass[10pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{babel}
\usepackage{fancybox}
\usepackage[unicode=true]{hyperref}
\usepackage{breakurl}

\usepackage[parfill]{parskip}

\begin{document}

\title{OpenSHMEM Tutorial Proposal for SuperComputing 2014}
\author{Tony Curtis <tonyc@uh.edu> \and Oscar Hernandez <oscar@ornl.gov> \and ...}

\maketitle

\section{Detailed Description}

Programming languages and libraries that leverage sophisticated
hardware capabilities have a natural advantage when used on today's
and tomorrow's high performance computer systems.  This tutorial
introduces the use of OpenSHMEM, a library API that can be used to
write programs that scale with a natural expression of 1-sided,
overlapped communication and computation.

The overarching goal is to present the capabilities of the OpenSHMEM
library to the audience with a view to driving the development of the
library through community involvement.  The bigger the user community,
the more input there will be to help decide future directions.

The OpenSHMEM project is an effort to develop the current API to make
it more expressive on current and future HPC systems.  The effort
includes development of the specification of OpenSHMEM, the web site,
tutorial, presentation, workshop and other outreach materials, along
with new tools that programmers can use to detect and avoid problems
in OpenSHMEM programs at compile-time.

People who attend this tutorial will find out about the history of
OpenSHMEM, its current state, and where we see it going.  Co-design of
software and hardware is a hot topic as we try to build ever larger
systems without running into power and financial brick walls.

Designing expressive programming environments that allow programmers
to solve problems is part of the solution, but also if we can closely
match the anticipated hardware capabilities, especially within network
fabrics, we can build yet more efficient systems.  Communication
layers that can take advantage of existing and novel network hardware
will allow the development of yet more interesting higher-level
environments such as libraries like OpenSHMEM and highly-productive
languages.  Oak Ridge National Laboratory has been developing one such
interface called the Uniform Common Communication Substrate (UCCS);
part of the tutorial will cover the current state of this library and
how we are working to provide a very high performing software
interface to networks.

Our intended audience are programmers and technology decision makers
who are looking for information to help them decide on future HPC
development methods.  The tutorial will be gradated to be attractive
to programmers who are unfamiliar with the details of OpenSHMEM and
PGAS environments, as well as more seasoned programmers and
technologists who want to acquaint themselves with the state of the
art.

The tutorial is intended for people who can program in C and/or
Fortran, or at least are familiar with such code, as those are the
languages that OpenSHMEM targets, but note that we will concentrate on
C for our examples.  A familiarity with parallel programming
environments will be be helpful, e.g. GNU/Linux usage, MPI, clusters
(job schedulers, parallel launchers).

\subsection{General Description}

The tutorial covers introductory material, going over the history of
OpenSHMEM to provide context for the current efforts, and discusses
the current state of play, before looking at complementary efforts for
network support and the tool ecosystem that will make OpenSHMEM a
highly productive environment.  We look at some example programs that
demonstrate how OpenSHMEM works, and will make available code
examples.  Installation and use of the library will be covered for
those who wish to participate in development.  Information about how
to propose new features and take part in online discussions will be
given so that attendees can continue to involve themselves in the
process after the tutorial.

\subsection{Cohesion of Presenters}

The tutorial will be presented by researchers from the University of
Houston (UH), and Oak Ridge National Laboratory (ORNL).  These 2
institutions have worked together for a number of years on a number of
projects.  As well as the 4-year OpenSHMEM effort, UH graduate
students have been perennial summer interns at ORNL over the past few
years, and there are UH post-doctorates working at ORNL, including one
of the presenters.  The development of the tutorial material thus
comes from a longstanding collaborative process, and we have given
joint tutorials and workshops in the past, including this year.

\subsection{Update Of Previous Material}

The material we intend to present has in part appeared before in
previous tutorials.  This is to aid in the familiarization process as
many attendees are new to OpenSHMEM.  To add to existing material, we
include some case studies of programs that were either ported from
other paradigms, e.g. MPI, or written from scratch, and show how to
progressively improve the parallelization afforded by OpenSHMEM.

\section{Detailed Outline}

\subsection{Overview}

  To set context, we will provide the audience with an outline of the
  space occupied by OpenSHMEM in the programming landscape, comparing
  it with other libraries and languages in the Partitioned Global
  Address Space family.  To show how the API is structured, we will
  take a guided tour of the API, highlighting with examples how
  OpenSHMEM is used in programs.

\subsection{The OpenSHMEM API}

  The API consists of 5 basic sections: memory management,
  point-to-point data transfers, synchronizations, collective
  operations, and atomics and locks.  Introduced in this order, we can
  incrementally describe working OpenSHMEM programs that grow in
  complexity.

\subsection{UCCS Development}

  We will discuss the current state and future plans of the UCCS
  substrate software.  Design decisions will be be covered with a view
  to explaining how the software is intended to be a platform not just
  for OpenSHMEM, but also for PGAS/HPCS languages.

\subsection{OpenSHMEM is evolving}

  We will present existing extension ideas (both at the time of
  writing and those in play at the time of the tutorial) and draw on
  the experience of the audience to actively solicit new ideas.

\subsection{Debugging and Profiling Tools}

  OpenSHMEM is supported by various well-known profiling and debugging
  tools, including TAU, Vampir, and DDT.  We will discuss their use
  briefly, but avoid in-depth discussion because other tutorials are
  likely to be focused on these topics specifically.

\subsection{Our new OpenSHMEM Analyzer tool}

  This will be covered to help attendees understand how to detect
  errors and inefficiencies in programs.  Examples will show how the
  tool is able to report on under-synchronized programs which can lead
  to race conditions, hangs or undefined behavior, and
  over-synchronized ones which introduce slowdowns.

\section{Hands-On Work}

The OpenSHMEM reference implementation and its dependencies are
available in source form for download.  The library includes demo
programs that attendees can examine and work with.  The software runs
both on a desktop/laptop and on larger clusters, giving a wide range
of deployment options for this tutorial.  For those interested in
obtaining the software themselves, we will provide assistance in
installing from source on GNU/Linux.  Hands-on sessions in this
tutorial will allow attendees to try things out at their own pace so
these activities can occur concurrently with the presenters available
to help.

We will walk attendees through installing the OpenSHMEM reference
implementation and GASNet, and then show how to use the library in a
number of configurations, including in a personal laptop and on a
cluster with Infiniband interconnect.  There will be discussion of
other common combinations.

The demonstration programs that come with the library, plus some other
sample programs are available for building and experimentation.
Having working examples helps people to see how to write their own
programs from scratch.

Some MPI programs with OpenSHMEM counterparts will be available to
show how existing parallel programs can be converted.

All tutorial material will be available for download.

\section{R\'{e}sum\'{e}s}

\subsection{Tony Curtis}

\subsubsection*{University of Houston, Computer Science Department:
  September 2013 - present}

    I am a senior member of Dr. Barbara Chapman's HPCTools research
    group in Computer Science.  My main work is lead on the OpenSHMEM
    project with 2 post-docs and 3 PhD students, as well as leading
    system administration of group resources.  I provide a consultant
    role in the university's High Performance Computing Center, of
    which Barbara Chapman is also director.

    For the OpenSHMEM project, we have presented workshops/tutorials
    at various conferences, for example PGAS11 and 12, and members of
    the team have presented posters and papers at other conferences.
    I have also led a number of OpenSHMEM BOFs at SC conferences.

\subsubsection*{HPCVL:
  February 2006 - November 2008}

    I am Operations Manager at HPCVL, one of the national level High
    Performance Computing Consortia in Canada, in Kingston, Ontario
    (Canada) at the Queen's University.

    Through managing a group of 4 system administrators, and
    interfacing with user support, security and web personnel, I am
    responsible for day-to-day operations and future planning of
    systems, services and infrastructure at the Kingston and satellite
    locations.

\subsubsection*{University of Houston High-Performance Computing:
  January 2000 - September 2005}

    I was the primary technical driver for a Sun Microsystems Center
    of Excellence SunFire cluster configuration and its implementation
    and I head the systems administration team.

    I helped guest teach some introductory HPC classes for the Computer
    Science department, e.g. MPI and Grid.

\subsubsection*{VCPC, European Centre for Parallel Computing:
  May 1996 - April 1999}

    As systems manager, I am responsible for the setup, running and
    maintenance of VCPC's machines (mainly Solaris and SGI
    workstations) plus overseeing the maintenance and running of the
    Meiko CS-2 service. This service is available to users world-wide.

    VCPC also performs consultancy to European SMEs and I have been
    involved directly in some projects and indirectly through platofrm
    support.

\subsubsection*{EUnet Austria in Vienna:
  May 1994 - April 1996}

    EUnet was a pan-European ISP, and in 1994 was expanding as
    commercial and private interest in the Internet and the WWW grew.
    My main work there was to set up the WWW server and to handle
    customers requiring WWW access and sites co-located on our server
    (virtual hosting etc.).  Also the work involved visiting customers
    and setting up on-site WWW / proxy servers, including servers for
    the Austrian Press Agency, Der Standard (broadsheet newspaper),
    and the federal Chamber of Commerce.

\subsubsection*{University of Kent at Canterbury:
  January 1988 - April 1994}

    At the University of Kent I was initially working in the parallel
    computing research team and also involved in undergraduate and
    postgraduate (MSc. course) teaching. The teaching included a
    number of topics, introduction to programming, theoretical
    computer science, logic programming, functional programming,
    mathematics, parallel computing and others.  The group also ran
    external paid courses in parallel programming for industry users.

\subsection{Oscar}

\subsection{Other}

\end{document}
